---
title: "Troy_Uninj_Inj. BiomaRt LandSCENT."
author: "Moa Stenudd"
date: "10/7/2021"
output: 
  html_document:
    theme: "cosmo"
    highlight: "zenburn"
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  tidy = TRUE,
  cache = TRUE,
  echo = FALSE
)
```

```{r load}
library(dplyr)
library(Seurat)
library(knitr)
library(ggplot2)
library(ggpubr)
library(patchwork)
library(RColorBrewer)
library(pheatmap)
library(biomaRt)
library(LandSCENT)
library(monocle)
library(scater)
```

## Load seurat object

```{r loadSeurat_Featurenames}
load('Troy_UI_Inj_after_analysis_pipe_TOM_regTom_2021_07_10.Robj')

#genesTroyUI <- read.table('~/Documents/CMB/10X/tom/10X_19_040_tom/outs/filtered_feature_bc_matrix/features.tsv', stringsAsFactors = FALSE)

#genesTroySCI <- read.table('~/Documents/CMB/10X/tom/10X_19_041_tom/outs/filtered_feature_bc_matrix/features.tsv', stringsAsFactors = FALSE)

# check if the two gene tables above are identical
#identical(genesTroyUI, genesTroySCI)

# They are, so I comment them out and make only one called "genes"

#genes <- read.table('~/Documents/CMB/10X/tom/10X_19_040_tom/outs/filtered_feature_bc_matrix/features.tsv', stringsAsFactors = FALSE)

```

```{r extractMatrix}
# Extract scaled data from seurat object
epc.scaled <- GetAssayData(object = epc.combined, slot = "scale.data")

# Extract normalised data from seurat object
epc.normalized <- as.matrix(GetAssayData(object = epc.combined, slot = "data"))

# Get the cell embeddings 
epc.embeddings <- Embeddings(object = epc.combined, reduction = "umap")
epc.clusters <- FetchData(object = epc.combined, vars = "seurat_clusters")
epc.line <- FetchData(object = epc.combined, vars = "line")

#Make column out of gene names (this is where you'll join the data to the annotation file)
epc.normalized <- as.data.frame(epc.normalized)
epc.normalized$MGI.symbol <- rownames(epc.normalized)
musList <- epc.normalized$MGI.symbol
```

# Changing from mouse to human gene annotation

## This seems to have generated some duplicates / some missed genes, I need to check if it is a  problem of certain genes only existing in 1 version in the human data  set while they exist in i e 5 in the mouse data set or vice versa - or if it is a problem with biomaRt. For now I will just remove duplicates so that I can try out the downstream processing, but they need to be handled better!

```{r biomaRt}
convertMouseGenes <- function(x){
require("biomaRt")
human <- useMart("ensembl", dataset = "hsapiens_gene_ensembl")
mouse <- useMart("ensembl", dataset = "mmusculus_gene_ensembl")

genesV2 <- getLDS(attributes = "mgi_symbol",
filters = "mgi_symbol",
values = x,
mart = mouse,
attributesL = "hgnc_symbol",
martL = human,
uniqueRows = TRUE)

return(genesV2)
}

annotation <- convertMouseGenes(musList)
```

# Join annotation with expression matrix

```{r joinClean}
# Join annotation and expression matrix by the mouse gene symbol
epc.normalized.annotated <- left_join(epc.normalized, annotation, by = "MGI.symbol")

# Remove rows (genes) that do not have an annotation in the HGNC column (human gene)
epc.normalized.annotated <- epc.normalized.annotated[!(is.na(epc.normalized.annotated$HGNC.symbol) | epc.normalized.annotated$HGNC.symbol==""), ]
```

# Check gene symbols

!!! Some duplicates found !!! Will be handled by deletion of duplicate entries.

```{r names}
move_to_start <- function(x, n) {
  x[, c(tail(seq_len(ncol(x)), n), seq_len(ncol(x) - n))]
} 

epc.normalized.annotated <- move_to_start(epc.normalized.annotated, 2)

# Check for duplicates, subset out any duplicates to a separate DF to check them easily
epc.anno.duplicates <- subset(epc.normalized.annotated, duplicated(epc.normalized.annotated$HGNC.symbol))
```

# Remove duplicates + set row names and remove gene name cols

```{r fixRownames}
#  Removing duplicates

# df = df[order(df[,'Date'],-df[,'Depth']),] <- sort based on something that you wanna keep by, so if you want to keep the one with the highest avg expression you make a col to calc avg expression and then sort by that col, and then remove with the row below

epc.normalized.annotated <- epc.normalized.annotated[!duplicated(epc.normalized.annotated$HGNC.symbol),]

# Make human gene names into row names
row.names(epc.normalized.annotated) <- epc.normalized.annotated$HGNC.symbol

# Remove gene name columns
epc.normalized.annotated <- subset(epc.normalized.annotated, select = -c(MGI.symbol, HGNC.symbol))
```

# Change row names to entrezID

```{r entrez}
require(AnnotationDbi)
require(org.Hs.eg.db)

anno.v <- mapIds(org.Hs.eg.db, keys = rownames(epc.normalized.annotated), keytype = "SYMBOL", 
                 column = "ENTREZID", multiVals = "first")
unique_anno.v <- unique(anno.v)
epc.normalized.annotated_New <- matrix(0, nrow = length(unique_anno.v), ncol = dim(epc.normalized.annotated)[2])

for (i in seq_len(length(unique_anno.v))) {
  tmp <- epc.normalized.annotated[which(anno.v == unique_anno.v[i]) ,]
  if (!is.null(dim(tmp))) {
    tmp <- colSums(tmp) / dim(tmp)[1]
  }
  epc.normalized.annotated_New[i ,] <- epc.normalized.annotated_New[i ,] + tmp
}
rownames(epc.normalized.annotated_New) <- unique_anno.v
colnames(epc.normalized.annotated_New) <- colnames(epc.normalized.annotated)
epc.normalized.annotated_New <- epc.normalized.annotated_New[-which(rownames(epc.normalized.annotated_New) %in% NA) ,]
epc.normalized.annotated <- epc.normalized.annotated_New
```

# Create single cell experiment

```{r singleCellExperiment}
library(SingleCellExperiment)

sce <- SingleCellExperiment(list(counts = epc.normalized.annotated),
   metadata=pData)
```

# Load PPI 
```{r PPI}
# Loading newer PPI from LandSCENT
data(net17Jan16.m)
```

```{r landSCENT}
Integration.l <- DoIntegPPI(exp.m = sce, ppiA.m = net17Jan16.m)
str(Integration.l)
 
SR.o <- CompSRana(Integration.l, local = TRUE, mc.cores = 8)
```

```{r SRplot, eval = False}
boxplot(SR.o$SR ~ SR.o$data.sce@metadata$line)

boxplot(SR.o$SR ~ SR.o$data.sce@metadata$seurat_clusters)
```

```{r combinePlotMetadata}
# SR from sce
SR_df <- as.data.frame(colData(SR.o$data.sce))
# Cluster assignments from seurat object
clusters_df <- as.data.frame(epc.combined$seurat_clusters)
# Line (condition) from seurat object
line_df <- as.data.frame(epc.combined$line)
# Umap coordinates from seurat object
umap_df <- as.data.frame(Embeddings(epc.combined[["umap"]]))

# Join into one df based on rowname
l <- list(SR_df, line_df, clusters_df, umap_df)
dfs2 <- Reduce(merge, lapply(l, function(x) data.frame(x, rn = row.names(x))))

save(dfs2, file = "SR_Troy_210710.robj")
```

```{r plot, fig.height = 3, fig.width = 3.5}
theme_set(theme_cowplot())

# Plot (identical to above just to check again that the data is good)
#boxplot(dfs$SR ~ dfs2$epc.combined.seurat_clusters)

ggplot(dfs2, aes(x = epc.combined.seurat_clusters, y = SR, col = epc.combined.line)) + geom_jitter() + xlab("Cluster") + scale_color_brewer("Condition", palette = "Dark2")
ggplot(dfs2, aes(x = epc.combined.seurat_clusters, y = SR)) + geom_boxplot() 
ggplot(dfs2, aes(x = epc.combined.line, y = SR)) + geom_boxplot() 

ggplot(dfs2, aes(x = UMAP_1, y = UMAP_2, col = epc.combined.seurat_clusters)) + geom_point() + scale_color_brewer("Cluster", palette = "Dark2")
ggplot(dfs2, aes(x = UMAP_1, y = UMAP_2, col = SR)) + geom_point() + scale_colour_gradientn(colours = c("lightgray", "yellow", "red"))

```

```{r violins, fig.height=3, fig.width=2.3}
median.stat <- function(x){
    out <- quantile(x, probs = c(0.5))
    names(out) <- c("ymed")
    return(out) 
}

ggplot(dfs2, 
       aes(x = epc.combined.seurat_clusters, 
           y = SR, 
           fill = epc.combined.seurat_clusters)) + 
  geom_violin(scale="width") + 
  xlab("Cluster") + 
  scale_fill_brewer(palette = "Dark2") +
  theme(legend.title=element_blank()) +
  stat_summary(fun = median.stat, geom = 'point', size = 3, colour = "black")

```

```{r loadSCE, eval = FALSE}
# Example from tutorial, not used
sce <- SingleCellExperiment(list(counts=counts),
    colData = DataFrame(label=pretend.cell.labels),
    rowData = DataFrame(length=pretend.gene.lengths),
    metadata = list(study="GSE111111")
)
```

